---
title: "UK Greenhouse Gas Inventory"
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#FFFFFF"
      fg: "#223022" 
      primary: "#96BC61"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: scroll
runtime: shiny
---

```{r load libraries, include=FALSE}

options(repos = list(CRAN="http://cran.rstudio.com/")) #tell to install all from CRAN

##Create groups of libraries
datalibraries <- c("dplyr", "tidyverse", "janitor", "readxl", "readODS") #data reading and  manipulation
vislibraries <- c("ggplot2", "ggthemes", "scales") #pretty plots
applibraries <- c("shiny", "plotly", "shinyWidgets", "flexdashboard", "DT") #libraries for app

##Install packages
#lapply(datalibraries, install.packages, character.only = TRUE)
#lapply(vislibraries, install.packages, character.only = TRUE)
#lapply(applibraries, install.packages, character.only = TRUE)

##Load libraries
lapply(datalibraries, library, character.only = TRUE)
lapply(vislibraries, library, character.only = TRUE)
lapply(applibraries, library, character.only = TRUE)

```

```{r global, include=FALSE}

##Map loop to download UK 2022 data from the UK Department for Energy Security and Net Zero
downloaded <- file.exists("UKGHG_2022.ods") #checks if file is downloaded in working directory
if(downloaded != T){ #if downloaded is not true
  map2("https://assets.publishing.service.gov.uk/media/65c0d54663a23d000dc821ca/final-greenhouse-gas-emissions-2022-by-source-dataset.ods", #update this link when new data available
       "UKGHG_2022.ods", download.file)} #else{print('data downloaded')} #name and download or print

##Read in ods file
GHG_UK22 <- read_ods(
  path = "UKGHG_2022.ods",
  sheet = 1, #define tab/sheet to read
  col_names = TRUE, #use header row for column names
  col_types = NULL, #guess data types
  na = "", #treat blank cells as NA
  skip = 0, #don't skip rows
  formula_as_formula = FALSE, #values only
  range = NULL, 
  row_names = FALSE, #no row names
  strings_as_factors = TRUE) %>% #use factors
  clean_names() %>% #clean column names to lowercase, with underscores
  separate_wider_delim(source, delim = " - ", #separate source, define delimiter as " - "
                       names = c("source_1", "source_2", "source_3"), #name new columns
                       too_few = "align_start", too_many = "merge", #what to do with fewer or more than 3
                       cols_remove = FALSE) %>% #keep original column
  separate_wider_delim(tes_category, delim = " - ", #separate category, define delimiter as " - "
                       names = c("category_1", "category_2"), #name new columns
                       too_few = "align_start", too_many = "merge", #what to do with fewer or more than 2
                       cols_remove = FALSE) %>% #keep original column
  separate_wider_delim(activity, delim = " - ", #separate activity, define delimiter as " - "
                       names = c("activity_1", "activity_2"), #name new columns
                       too_few = "align_start", too_many = "merge", #what to do with fewer or more than 2
                       cols_remove = FALSE) %>% #keep original column
  replace(is.na(.), "") #get rid of NA values, input blank

#Define custom color palette ####
colors <- c(
  "#3CBD9C", #transport
           "#F0A7CA", #buildings
           "#F0C954", #industry
           "#C24841", #electricity
           "#96BC61", #ag green
           "#5D7ED1", #lulucf blue
           "#D38744", #fuel
           "#A780BB"  #waste purple
)


colors_3lvls <- c(
  "#C24841", #CH4
           "#3CBD9C", #CO2
           "#F0C954"  #N2O
)

#Set new error message across all errors from input
if (!interactive()) {
  options(shiny.error = function() {
    stop("Insufficient data. Please change your selection on the left.")
  })
}

```

Sidebar {.sidebar}
======================================================================


### Filters

```{r}

pickerInput("select_GHG", #input name
            "GHG group", #display text
            choices = c("CO2","CH4","N2O","NF3","HFCs","PFCs","SF6"), 
            multiple = TRUE,
            options = list(`actions-box` = TRUE), #creates buttons
            selected = c("CO2","CH4","N2O","NF3","HFCs","PFCs","SF6") #initial values 
            )


sliderInput("select_Year", #input name
            label = "Year:", #display text
            min = 1991, max = 2022, 
            value = 2022, #initial value
            step = 1,
            sep = "") #remove commas


```

Select gases and year you would like to view to update the figures to the right.

Emissions are reported in megatonnes (Mt) of carbon dioxide (CO2) equivalents (eq.). 

Data from the [Department for Energy Security and Net Zero](https://www.data.gov.uk/dataset/9568363e-57e5-4c33-9e00-31dc528fcc5a/uk-greenhouse-gas-emissions-final).

High-Level View Across All Sectors {data-orientation=rows}
=======================================================================
Row
-----------------------------------------------------------------------
### GHG Emissions (Mt CO2 eq.)

```{r, error=FALSE, out.width = '70%'}

renderValueBox({

#Calculate total positive emissions
emissions <- GHG_UK22 %>%
  filter(emissions_mt_co2e >= "0") %>%
  filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
  summarize(sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  round(., 1) #1 digit after decimal

valueBox(emissions, 
         icon = "ion-fireball", color = "#C24841")

})


```

### Negative Emissions (Mt CO2 eq.)

```{r, error=FALSE,}

renderValueBox({
  
#Calculate total negative emissions
removals <- GHG_UK22 %>%
  filter(emissions_mt_co2e <= "0") %>%
  filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
  summarize(sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  round(., 1) #1 digit after decimal

valueBox(removals,
         icon = "ion-leaf", color = "#3CBD9C")

})

```

### Net Emissions (Mt CO2 eq.)

```{r, error=FALSE,}

renderValueBox({
  
#Calculate total emissions
net <- GHG_UK22 %>%
  filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
  summarize(sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  round(., 1) #1 digit after decimal

valueBox(net, 
         icon = "ion-load-a", color = "#F0C954")

})

```

Row {data-height=650}
-----------------------------------------------------------------------
### Greenhouse Gas (GHG) Emissions by Sector
Click a division within the pie to "zoom in" and view a further breakdown of that sector or category. Click the center to "zoom out" back towards the broader view. 
```{r, error=FALSE}

produce_eplot <- shiny::reactive({
  
  ## Emissions data format for sunburst ####
  
  ## Apply filters
  GHG_pos <- GHG_UK22 %>%
    filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
    filter(emissions_mt_co2e >= "0") #subset filter only positive values
  
  #Run validate an input is selected
  validate(
  need(input$select_GHG != "", "Insufficient data. Please change your selection on the left.")
  )
  
  ##Separate out variables summaries
  sectors <- GHG_pos %>%
    group_by(tes_sector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE))
  
  subsectors <- GHG_pos %>%
    group_by(tes_sector, tes_subsector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE))
  
  cat_2 <- GHG_pos %>%
    filter(category_2 != "") %>%
    group_by(tes_sector, tes_subsector, category_2) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    unite('parents', c(tes_sector, tes_subsector), remove = TRUE, sep = " - ") #combine higher categories into parents
  
  cat_1 <- GHG_pos %>%
    group_by(tes_sector, tes_subsector, category_2, category_1) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    mutate(parents = case_when(
      category_2 != "" ~str_c(tes_sector,tes_subsector,category_2, sep = " - "),
      category_2 == "" ~str_c(tes_sector,tes_subsector, sep = " - ")))
  
  act_1 <- GHG_pos %>%
    group_by(tes_sector, tes_subsector, category_2, category_1, activity_1) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    mutate(parents = case_when(
      category_2 != "" ~str_c(tes_sector,tes_subsector,category_2, category_1, sep = " - "),
      category_2 == "" ~str_c(tes_sector,tes_subsector, category_1, sep = " - ")
    ))
  
  act_2 <- GHG_pos %>%
    filter(activity_2 != "") %>%
    group_by(tes_sector, tes_subsector, category_2, category_1, activity_1, activity_2) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    mutate(parents = case_when(
      category_2 != "" ~str_c(tes_sector,tes_subsector,category_2, category_1, activity_1, sep = " - "),
      category_2 == "" ~str_c(tes_sector,tes_subsector, category_1, activity_1, sep = " - ")
    ))
  
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - start simple, set root
  GHGsb1 <- data.frame(
    parents =c(rep("", 8)),
    labels =c(paste0(sectors$tes_sector)),
    values =as.numeric(c(paste0(sectors$Temissions)))
  )
  
  ##Add id column
  GHGsb1 <- GHGsb1 %>%
    mutate(ids = labels) 
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - add on
  GHGsb2 <- data.frame(
    parents =c(paste0(subsectors$tes_sector), paste0(cat_2$parents), paste0(cat_1$parents), 
               paste0(act_1$parents), paste0(act_2$parents)),
    labels =c(paste0(subsectors$tes_subsector), paste0(cat_2$category_2), paste0(cat_1$category_1), 
              paste0(act_1$activity_1), paste0(act_2$activity_2)),
    values =as.numeric(c(paste0(subsectors$Temissions), paste0(cat_2$Temissions), paste0(cat_1$Temissions), paste0(act_1$Temissions), paste0(act_2$Temissions)))
  )
  
  ##Add id column                    
  GHGsb2 <- GHGsb2 %>%
    unite('ids', c(parents, labels), remove = FALSE, sep = " - ") #combine columns for id
  
  ##Combine into one dataframe
  GHGsb <- rbind(GHGsb1, GHGsb2) 
  
  
  ## Sunburst Emissions with Plotly ####
  sb <- plot_ly(GHGsb,
                type = 'sunburst',
                ids = ~ids,
                labels = ~labels,
                parents = ~parents,
                values = ~values,
                branchvalues = 'total', #parents are totals of children
                insidetextorientation='radial',
                maxdepth = 2, #how many levels to show
                level = "",
                sort = TRUE #sort by size #plot dimensions
  ) %>%
    layout(
      #margin = list(l = 10, r = 10, b = 10, t = 10),
      sunburstcolorway = colors,
      extendsunburstcolors = TRUE
    ) #%>%
    style(sb, hovertemplate="%{label} 
      %{value:,.3r} Mt CO2 equivalents
      %{percentParent:.1%} of %{parent} emissions
      <extra></extra>")
  
  #sb
  
  })

renderPlotly({produce_eplot()})

```

Row
-----------------------------------------------------------------------
### Negative CO2 Emissions (Removals)

```{r, error=FALSE,}

produce_rplot <- shiny::reactive({
  
  ## Removals data format for sunburst ####
  
  ##Apply filters
GHG_neg <- GHG_UK22 %>%
  filter(year == input$select_Year) %>%
  filter(emissions_mt_co2e <= "0") #subset filter only negative values

  ##Replace empty source 2 values with source 1
  GHG_neg$source_2[GHG_neg$source_2 == ""] <- NA #also works to replace blanks or any value defined here with NA across a dataframe
  GHG_neg$source_2[is.na(GHG_neg$source_2)] <- GHG_neg$source_1[is.na(GHG_neg$source_2)]
  
  ##Separate out variables summaries - all GHGs combined
  subsectors_neg <- GHG_neg %>%
    group_by(tes_subsector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE))
  
  cat_neg <- GHG_neg %>%
    group_by(tes_subsector, tes_category) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) 
  
  src_neg <- GHG_neg %>%
    group_by(tes_subsector, tes_category, source_2) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    unite('parents', c(tes_subsector, tes_category), remove = TRUE, sep = " - ") #combine higher categories into parents
  
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - start simple, set root
  GHGsb1_neg <- data.frame(
    parents =c(rep("", 6)),
    labels =c(paste0(subsectors_neg$tes_subsector)),
    values =as.numeric(c(paste0(subsectors_neg$Temissions)))
  )
  
  ##Add id column
  GHGsb1_neg <- GHGsb1_neg %>%
    mutate(ids = labels) 
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - add on
  GHGsb2_neg <- data.frame(
    parents =c(paste0(cat_neg$tes_subsector), paste0(src_neg$parents)),
    labels =c(paste0(cat_neg$tes_category), paste0(src_neg$source_2)),
    values =as.numeric(c(paste0(cat_neg$Temissions), paste0(src_neg$Temissions)))
  )
  
  ##Add id column                    
  GHGsb2_neg <- GHGsb2_neg %>%
    unite('ids', c(parents, labels), remove = FALSE, sep = " - ") #combine columns for id
  
  ##Combine into one dataframe
  GHGsb_neg <- rbind(GHGsb1_neg, GHGsb2_neg) 
  
  ##Convert negative values to absolute (positive) values
  GHGsb_neg$values <- abs(GHGsb_neg$values)
  
  ## Sunburst Removals with Plotly ####
  sb_neg <- plot_ly(GHGsb_neg,
                    type = 'sunburst',
                    ids = ~ids,
                    labels = ~labels,
                    parents = ~parents,
                    values = ~values,
                    branchvalues = 'total', #parents are totals of children
                    insidetextorientation='radial',
                    maxdepth = 2, #how many levels to show
                    level = "",
                    sort = TRUE #sort by size
                    #width = 800, height = 800 #plot dimensions
  ) %>%
    layout(
      #margin = list(l = 10, r = 10, b = 10, t = 10),
      sunburstcolorway = c(
        "#3CBD9C", #forestry
                 "#F0C954", #grassland
                 "#5D7ED1", #peatland
                 "#C24841", #other
                 "#96BC61",
                 "#F0A7CA", 
                 "#D38744", 
                 "#A780BB"
      ),
      extendsunburstcolors = TRUE
    ) #%>%
    style(sb_neg, hovertemplate="%{label} 
      %{value:,.3r} Mt CO2 equivalents
      %{percentParent:.1%} of %{parent} removals
      <extra></extra>")
  
  #sb_neg
  })

plotly::renderPlotly({produce_rplot()})

```
### Time Series

```{r, error=FALSE,}

produce_tplot <- shiny::reactive({
  
  #Run validate an input is selected
  validate(
  need(input$select_GHG != "", "Insufficient data. Please change your selection on the left.")
  )
  
  #Plot time series ggplot
  Sect_time <- GHG_UK22 %>%
    filter(ghg_grouped %in% input$select_GHG) %>%
    group_by(year, tes_sector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    rename("Sector" = "tes_sector") %>% #rename variable
    ggplot(aes(x=as.numeric(year), y=Temissions, color=Sector)) +
    geom_line() +
    xlab("") + ylab("Greenhouse Gas Emissions (Mt CO2 eq.)") +
    ggtitle("Emissions per Sector Over Time") +
    theme_few() +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = pretty(c(1990:2020), n=6)) +
    scale_color_manual(values = c("Agriculture" = "#96BC61", 
                                 "Buildings and product uses" = "#F0A7CA", 
                                 "Domestic transport" = "#3CBD9C", 
                                 "Electricity supply" = "#C24841",
                                 "Fuel supply" = "#D38744", 
                                 "Industry" = "#F0C954", 
                                 "LULUCF" = "#5D7ED1", 
                                 "Waste" = "#A780BB"  
           )) +
    theme(legend.position="bottom")
  
  #Sect_time
    
    #Make interactive with Plotly
    t_int <- plotly::ggplotly(Sect_time) %>% #convert to interactive graph
      layout(legend = list(
        orientation = "h")) #%>%
    style(t_int, hovertemplate="Year: %{x:,.r}
          CO2eq: %{y:,.4r} Mt") #define what hover label says, round x values, round y values to 4 sig. figs
    
    #t_int
  
  })


renderPlotly({produce_tplot()})


```

Row
-----------------------------------------------------------------------
### Summary Emissions Per Sector

```{r}

DT::renderDataTable({
  
  #Create filtered dataframe with columns for positive and negative emissions
  GHG_table <- GHG_UK22 %>%
    filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
    mutate(pos_e = ifelse(emissions_mt_co2e >= 0, emissions_mt_co2e, 0)) %>%
    mutate(neg_e = ifelse(emissions_mt_co2e <= 0, emissions_mt_co2e, 0)) %>%
    rename("Sector" = "tes_sector") #%>% #rename variable
    #rename("Subsector" = "tes_subsector") #rename variable
  
  #Summarize positive emissions
  GHG_postable <- GHG_table %>%
    #group_by(Sector, Subsector) %>%
    group_by(Sector) %>%
    summarize("Positive Emissions" = sum(pos_e, na.rm = TRUE))
  
  #Summarize negative emissions
  GHG_negtable <- GHG_table %>%
    #group_by(Sector, Subsector) %>%
    group_by(Sector) %>%
    summarize("Negative Emissions" = sum(neg_e, na.rm = TRUE))
  
  #Combine positive and negative into one table
  GHG_posneg <- inner_join(GHG_postable, GHG_negtable,
                       by= c('Sector')) 
  
  #Summarize net emissions
  GHG_nettable <- GHG_table %>%
    #group_by(Sector, Subsector) %>%
    group_by(Sector) %>%
    summarize("Net Emissions" = sum(emissions_mt_co2e, na.rm = TRUE))
  
  #Combine into one summary table
  GHG_summary <- inner_join(GHG_posneg, GHG_nettable,
                       by= c('Sector')) 
  
  #Round numbers
  GHG_summary$`Positive Emissions` <- round(GHG_summary$`Positive Emissions`, 1)
  GHG_summary$`Negative Emissions` <- round(GHG_summary$`Negative Emissions`, 1)
  GHG_summary$`Net Emissions` <- round(GHG_summary$`Net Emissions`, 1)
  
  #Define number of rows to show
  summary_rows <- nrow(GHG_summary)
  GHG_summary <- head(GHG_summary, n = summary_rows)
  
  #Call Datatable, options
  DT::datatable(GHG_summary, 
                rownames = FALSE, #no row number labels
                options = list(#autoWidth = TRUE, #scale to element/window size
                               dom = 't', #hide search box
                               bPaginate = FALSE #don't need more pages
  ))
  
})

```


Defra Sector Graphs {data-orientation=rows}
=======================================================================

Row {.tabset .tabset-fade data-height=5000}
-------------------------------------
   
### Agriculture 

#### Agriculture Emissions by Subsector and GHG

```{r, error=FALSE,}

produce_agplot <- shiny::reactive({

#Bar plot by subsector, stacked by GHG with Plotly
agbar_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(year == input$select_Year) %>%
  group_by(tes_subsector, ghg, ghg_grouped) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = ghg_grouped, values_from=Temissions) %>%
  droplevels() %>%
  plot_ly(x = ~tes_subsector, type = 'bar', y = ~CO2, name = 'CO2', marker = list(color ="#3CBD9C")) %>%
  add_trace(y = ~CH4, name = 'CH4', marker = list(color ="#C24841")) %>%
  add_trace(y = ~N2O, name = 'N2O', marker = list(color ="#D38744")) %>%
  layout(yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         barmode = 'stack'
         )

#Plotly customization
style(agbar_int, hovertemplate="Subsector: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agplot()})


```   

#### Agriculture Emissions by Category

```{r, error=FALSE}

produce_agcatplot <- shiny::reactive({
  
#Bar plot by category, filled by subsector with Plotly
agcat_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
  group_by(tes_subsector, tes_category) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  droplevels() %>%
  plot_ly(x = ~tes_category, type = 'bar', y = ~Temissions, color = ~tes_subsector, colors = colors,
          marker = list(
            width = 1.5)) %>%
  layout(title = 'Agriculture Emissions by TES Subsector and Category',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'),
         xaxis = list(title = '', tickangle = -90),
         legend = list(x = 0.8, y = 1, orientation = 'v'),
         autosize = F, width = 1000, height = 600,
         autoexpand = T,
         barmode = 'group'
         )

#Plotly customization
style(agcat_int, 
hovertemplate="Category: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agcatplot()})
  
```

#### Agricultural Combustion

```{r}

produce_agcom <- shiny::reactive({
  
#Bar plot combustion, stacked by fuel type with Plotly
agcom_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(tes_subsector == "Agricultural combustion") %>%
  filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
  group_by(category_1, activity) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = activity, values_from=Temissions) %>%
  clean_names() %>%
  droplevels() %>%
  plot_ly(x = ~category_1, type = 'bar', y = ~gas_oil, name = 'Gas oil', marker = list(color ="#C24841")) %>%
  add_trace(y = ~petrol, name = 'Petrol', marker = list(color ="#A780BB")) %>%
  add_trace(y = ~biogas, name = 'Biogas', marker = list(color ="#3CBD9C")) %>%
  add_trace(y = ~burning_oil, name = 'Burning oil', marker = list(color ="#F0A7CA")) %>%
  add_trace(y = ~fuel_oil, name = 'Fuel oil', marker = list(color ="#D38744")) %>%
  add_trace(y = ~natural_gas, name = 'Natural gas', marker = list(color ="#5D7ED1")) %>%
  add_trace(y = ~straw, name = 'Straw', marker = list(color ="#F0C954")) %>%
  layout(title = 'Agricultural Combustion Emissions by Fuel Type',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         xaxis = list(title = ''),
         barmode = 'stack'
         )

#Plotly customization
style(agcom_int, 
hovertemplate="Source: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agcom()})

```

#### Agricultural Soils

```{r}

produce_agsoil <- shiny::reactive({

#Bar plot soil, stacked by GHG with Plotly
agsoil_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(tes_subsector == "Agricultural soils") %>%
  filter(year == input$select_Year) %>%
  group_by(tes_category, ghg) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = ghg, values_from=Temissions) %>%
  droplevels() %>%
  plot_ly(x = ~tes_category, type = 'bar', y = ~CO2, name = 'CO2', marker = list(color ="#3CBD9C")) %>%
  add_trace(y = ~N2O, name = 'N2O', marker = list(color ="#D38744")) %>%
  layout(title = 'Agricultural Soil Emissions by Category',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         xaxis = list(title = ''), 
         barmode = 'stack'
         )
  
#Plotly customization
style(agsoil_int, 
hovertemplate="Source: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agsoil()})

```

#### Livestock Emissions by Category and GHG

```{r}

produce_agls1 <- shiny::reactive({

#Bar plot livestock animals, stacked by GHG with Plotly
agls_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(tes_subsector == "Livestock") %>%
  filter(year == input$select_Year) %>%
  group_by(category_1, ghg) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = ghg, values_from=Temissions) %>%
  droplevels() %>%
  plot_ly(x = ~category_1, type = 'bar', y = ~CH4, name = 'CH4', marker = list(color ="#C24841")) %>%
  add_trace(y = ~N2O, name = 'N2O', marker = list(color ="#D38744")) %>%
  layout(title = 'Agricultural Livestock Emissions by Category',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         xaxis = list(title = ''), 
         barmode = 'stack'
         )
  
#Plotly customization
style(agls_int, 
hovertemplate="Source: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agls1()})

```

#### Livestock Emissions by Animal and GHG

```{r}

produce_agls2 <- shiny::reactive({
  
#Bar plot livestock animals, stacked by GHG with Plotly
agls2_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(tes_subsector == "Livestock") %>%
  filter(year == input$select_Year) %>%
  group_by(category_2, ghg) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = ghg, values_from=Temissions) %>%
  droplevels() %>%
  plot_ly(x = ~category_2, type = 'bar', y = ~CH4, name = 'CH4', marker = list(color ="#C24841")) %>%
  add_trace(y = ~N2O, name = 'N2O', marker = list(color ="#D38744")) %>%
  layout(title = 'Agricultural Livestock Emissions by Animal',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         xaxis = list(title = ''), 
         barmode = 'stack'
         )
  
#Plotly customization
style(agls2_int, 
hovertemplate="Source: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agls2()})

```

#### Livestock Emissions by Category and Animal

```{r}

produce_agls3 <- shiny::reactive({

#Bar plot livestock category, stacked by animal with Plotly
agls3_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(tes_subsector == "Livestock") %>%
  filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
  group_by(category_1, category_2) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = category_2, values_from=Temissions) %>%
  clean_names() %>%
  droplevels() %>%
  plot_ly(x = ~category_1, type = 'bar', y = ~beef_cattle, name = 'Beef cattle', marker = list(color ="#C24841")) %>%
  add_trace(y = ~dairy_cattle, name = 'Dairy cattle', marker = list(color ="#D38744")) %>%
  add_trace(y = ~sheep, name = 'Sheep', marker = list(color ="#96BC61")) %>%
  add_trace(y = ~pigs, name = 'Pigs', marker = list(color ="#F0A7CA")) %>%
  add_trace(y = ~poultry, name = 'Poultry', marker = list(color ="#A780BB")) %>%
  add_trace(y = ~other_livestock, name = 'Other livestock', marker = list(color ="#3CBD9C")) %>%
  layout(title = 'Agricultural Livestock Emissions by Category and Animal',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         xaxis = list(title = ''), 
         barmode = 'stack'
         )
  
#Plotly customization
style(agls3_int, 
hovertemplate="Source: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agls3()})

```

#### Other Agriculture Emissions

```{r}

produce_agother <- shiny::reactive({
  
#Bar plot other ag, stacked by GHG
agother_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  filter(tes_subsector == "Other agriculture") %>%
  filter(year == input$select_Year) %>%
  group_by(tes_category, ghg) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = ghg, values_from=Temissions) %>%
  droplevels() %>%
  plot_ly(x = ~tes_category, type = 'bar', y = ~CO2, name = 'CO2', marker = list(color ="#3CBD9C")) %>%
  layout(title = 'Other Agricultural Emissions',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         xaxis = list(title = ''), 
         barmode = 'stack'
         )
  
#Plotly customization
style(agother_int, 
hovertemplate="Source: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agother()})

```

#### Summary Emissions Per Subsector

```{r}

DT::renderDataTable({
  
  #Create filtered dataframe with columns for positive and negative emissions
  ag_table <- GHG_UK22 %>%
    filter(tes_sector == "Agriculture") %>%
    filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
    mutate(pos_e = ifelse(emissions_mt_co2e >= 0, emissions_mt_co2e, 0)) %>%
    mutate(neg_e = ifelse(emissions_mt_co2e <= 0, emissions_mt_co2e, 0)) %>%
    rename("Subsector" = "tes_subsector") #rename variable
  
  #Summarize positive emissions
  ag_postable <- ag_table %>%
    group_by(Subsector, category_2, category_1, activity_1, activity_2) %>%
    summarize("Positive Emissions" = sum(pos_e, na.rm = TRUE))
  
  #Summarize negative emissions
  ag_negtable <- ag_table %>%
    group_by(Subsector, category_2, category_1, activity_1, activity_2) %>%
    summarize("Negative Emissions" = sum(neg_e, na.rm = TRUE))
  
  #Combine positive and negative into one table
  ag_posneg <- inner_join(ag_postable, ag_negtable,
                       by= c('Subsector','category_2','category_1','activity_1','activity_2')) 
  
  #Summarize net emissions
  ag_nettable <- ag_table %>%
    group_by(Subsector, category_2, category_1, activity_1, activity_2) %>%
    summarize("Net Emissions" = sum(emissions_mt_co2e, na.rm = TRUE))
  
  #Combine into one summary table
  ag_summary <- inner_join(ag_posneg, ag_nettable,
                       by= c('Subsector','category_2','category_1','activity_1','activity_2')) 
  
  #Round numbers
  ag_summary$`Positive Emissions` <- round(ag_summary$`Positive Emissions`, 1)
  ag_summary$`Negative Emissions` <- round(ag_summary$`Negative Emissions`, 1)
  ag_summary$`Net Emissions` <- round(ag_summary$`Net Emissions`, 1)
  
  #Define number of rows to show
  agsummary_rows <- nrow(ag_summary)
  ag_summary <- head(ag_summary, n = agsummary_rows)
  
  #Call Datatable, options
  DT::datatable(ag_summary, 
                rownames = FALSE, #no row number labels
                options = list(#autoWidth = TRUE, #scale to element/window size
                               #dom = 't', #hide search box
                               bPaginate = FALSE #don't need more pages
  ))
  
})

```


### Waste

```{r}


```   


### Land Use, Land Use Change, and Forestry

```{r}

produce_lulucf <- shiny::reactive({

})

renderPlotly({produce_lulucf()})

```   


