---
title: "UK Greenhouse Gas Inventory"
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#101010"
      fg: "#FDF7F7" 
      primary: "#ED79F9"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: scroll
runtime: shiny
---

```{r load libraries, include=FALSE}
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()

##Create groups of libraries
datalibraries <- c("dplyr", "tidyverse", "janitor", "readxl", "readODS") #data reading and  manipulation
vislibraries <- c("ggplot2", "ggthemes", "RColorBrewer", "scales") #pretty plots
applibraries <- c("shiny", "plotly", "gridlayout", "bslib", "DT", "shinyWidgets") #libraries for app

##Install packages
#lapply(datalibraries, install.packages, character.only = TRUE)
#lapply(vislibraries, install.packages, character.only = TRUE)
#lapply(applibraries, install.packages, character.only = TRUE)
#devtools::install_github("rstudio/gridlayout") #install from github for this version of R

##Load libraries
lapply(datalibraries, library, character.only = TRUE)
lapply(vislibraries, library, character.only = TRUE)
lapply(applibraries, library, character.only = TRUE)
```

```{r global, include=FALSE}

##Map loop to download UK 2022 data from the UK Department for Energy Security and Net Zero
downloaded <- file.exists("UKGHG_2022.ods") #checks if file is downloaded in working directory
if(downloaded != T){ #if downloaded is not true
  map2("https://assets.publishing.service.gov.uk/media/65c0d54663a23d000dc821ca/final-greenhouse-gas-emissions-2022-by-source-dataset.ods", #update this link when new data available
       "UKGHG_2022.ods", download.file)} #else{print('data downloaded')} #name and download or print

##Read in ods file
GHG_UK22 <- read_ods(
  path = "UKGHG_2022.ods",
  sheet = 1, #define tab/sheet to read
  col_names = TRUE, #use header row for column names
  col_types = NULL, #guess data types
  na = "", #treat blank cells as NA
  skip = 0, #don't skip rows
  formula_as_formula = FALSE, #values only
  range = NULL, 
  row_names = FALSE, #no row names
  strings_as_factors = TRUE) %>% #use factors
  clean_names() %>% #clean column names to lowercase, with underscores
  separate_wider_delim(source, delim = " - ", #separate source, define delimiter as " - "
                       names = c("source_1", "source_2", "source_3"), #name new columns
                       too_few = "align_start", too_many = "merge", #what to do with fewer or more than 3
                       cols_remove = FALSE) %>% #keep original column
  separate_wider_delim(tes_category, delim = " - ", #separate category, define delimiter as " - "
                       names = c("category_1", "category_2"), #name new columns
                       too_few = "align_start", too_many = "merge", #what to do with fewer or more than 2
                       cols_remove = FALSE) %>% #keep original column
  separate_wider_delim(activity, delim = " - ", #separate activity, define delimiter as " - "
                       names = c("activity_1", "activity_2"), #name new columns
                       too_few = "align_start", too_many = "merge", #what to do with fewer or more than 2
                       cols_remove = FALSE) %>% #keep original column
  replace(is.na(.), "") #get rid of NA values, input blank

#Define custom color palette ####
colors <- c(
  "#3CBD9C", #transport
           "#F0A7CA", #buildings
           "#F0C954", #industry
           "#C24841", #electricity
           "#96BC61", #ag green
           "#5D7ED1", #lulucf blue
           "#D38744", #fuel
           "#A780BB"  #waste purple
)


colors_3lvls <- c(
  "#C24841", #CH4
           "#3CBD9C", #CO2
           "#F0C954"  #N2O
)


```

Sidebar {.sidebar}
======================================================================


### Filters

```{r}

selectInput(inputId = "select_GHG", #input name
            label = "GHG group", #display text
            choices = c("CO2","CH4","N2O","NF3","HFCs","PFCs","SF6"), 
            multiple = TRUE,
            selected = c("CO2","CH4","N2O"))


sliderInput("select_Year", #input name
            label = "Year:", #display text
            min = 1991, max = 2022, 
            value = 2022, #initial value
            step = 1,
            sep = "") #remove commas


```

The UK's greenhouse gas (GHG) inventory background...

Emissions are reported in megatonnes (Mt) of carbon dioxide (CO2) equivalents (eq.). 



Created by Kaydee Barker for Defra's Net Zero Teams.


High-Level View Across All Sectors
=======================================================================


Row
-----------------------------------------------------------------------

### Greenhouse Gas (GHG) Emissions by Sector

```{r}

produce_eplot <- shiny::reactive({
  
  ## Emissions data format for sunburst ####
  
  ## Apply filters
  GHG_pos <- GHG_UK22 %>%
    filter(year == input$select_Year & ghg_grouped %in% input$select_GHG) %>%
    filter(emissions_mt_co2e >= "0") #subset filter only positive values
  
  ##Separate out variables summaries
  sectors <- GHG_pos %>%
    group_by(tes_sector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE))
  
  subsectors <- GHG_pos %>%
    group_by(tes_sector, tes_subsector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE))
  
  cat_2 <- GHG_pos %>%
    filter(category_2 != "") %>%
    group_by(tes_sector, tes_subsector, category_2) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    unite('parents', c(tes_sector, tes_subsector), remove = TRUE, sep = " - ") #combine higher categories into parents
  
  cat_1 <- GHG_pos %>%
    group_by(tes_sector, tes_subsector, category_2, category_1) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    mutate(parents = case_when(
      category_2 != "" ~str_c(tes_sector,tes_subsector,category_2, sep = " - "),
      category_2 == "" ~str_c(tes_sector,tes_subsector, sep = " - ")))
  
  act_1 <- GHG_pos %>%
    group_by(tes_sector, tes_subsector, category_2, category_1, activity_1) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    mutate(parents = case_when(
      category_2 != "" ~str_c(tes_sector,tes_subsector,category_2, category_1, sep = " - "),
      category_2 == "" ~str_c(tes_sector,tes_subsector, category_1, sep = " - ")
    ))
  
  act_2 <- GHG_pos %>%
    filter(activity_2 != "") %>%
    group_by(tes_sector, tes_subsector, category_2, category_1, activity_1, activity_2) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    mutate(parents = case_when(
      category_2 != "" ~str_c(tes_sector,tes_subsector,category_2, category_1, activity_1, sep = " - "),
      category_2 == "" ~str_c(tes_sector,tes_subsector, category_1, activity_1, sep = " - ")
    ))
  
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - start simple, set root
  GHGsb1 <- data.frame(
    parents =c(rep("", 8)),
    labels =c(paste0(sectors$tes_sector)),
    values =as.numeric(c(paste0(sectors$Temissions)))
  )
  
  ##Add id column
  GHGsb1 <- GHGsb1 %>%
    mutate(ids = labels) 
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - add on
  GHGsb2 <- data.frame(
    parents =c(paste0(subsectors$tes_sector), paste0(cat_2$parents), paste0(cat_1$parents), 
               paste0(act_1$parents), paste0(act_2$parents)),
    labels =c(paste0(subsectors$tes_subsector), paste0(cat_2$category_2), paste0(cat_1$category_1), 
              paste0(act_1$activity_1), paste0(act_2$activity_2)),
    values =as.numeric(c(paste0(subsectors$Temissions), paste0(cat_2$Temissions), paste0(cat_1$Temissions), paste0(act_1$Temissions), paste0(act_2$Temissions)))
  )
  
  ##Add id column                    
  GHGsb2 <- GHGsb2 %>%
    unite('ids', c(parents, labels), remove = FALSE, sep = " - ") #combine columns for id
  
  ##Combine into one dataframe
  GHGsb <- rbind(GHGsb1, GHGsb2) 
  
  
  ## Sunburst Emissions with Plotly ####
  sb <- plot_ly(GHGsb,
                type = 'sunburst',
                ids = ~ids,
                labels = ~labels,
                parents = ~parents,
                values = ~values,
                branchvalues = 'total', #parents are totals of children
                insidetextorientation='radial',
                maxdepth = 2, #how many levels to show
                level = "",
                sort = TRUE #sort by size #plot dimensions
  ) %>%
    layout(
      #margin = list(l = 10, r = 10, b = 10, t = 10),
      sunburstcolorway = colors,
      extendsunburstcolors = TRUE
    ) #%>%
    style(sb, hovertemplate="%{label} 
      %{value:,.3r} Mt CO2 equivalents
      %{percentParent:.1%} of %{parent} emissions
      <extra></extra>")
  
  #sb
  
  })

renderPlotly({produce_eplot()})

```

Row
-----------------------------------------------------------------------

### Negative CO2 Emissions (Removals)

```{r}

produce_rplot <- shiny::reactive({
  
  ## Removals data format for sunburst ####
  
  ##Apply filters
GHG_neg <- GHG_UK22 %>%
  filter(year == input$select_Year) %>%
  filter(emissions_mt_co2e <= "0") #subset filter only negative values

  ##Replace empty source 2 values with source 1
  GHG_neg$source_2[GHG_neg$source_2 == ""] <- NA #also works to replace blanks or any value defined here with NA across a dataframe
  GHG_neg$source_2[is.na(GHG_neg$source_2)] <- GHG_neg$source_1[is.na(GHG_neg$source_2)]
  
  ##Separate out variables summaries - all GHGs combined
  subsectors_neg <- GHG_neg %>%
    group_by(tes_subsector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE))
  
  cat_neg <- GHG_neg %>%
    group_by(tes_subsector, tes_category) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) 
  
  src_neg <- GHG_neg %>%
    group_by(tes_subsector, tes_category, source_2) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    unite('parents', c(tes_subsector, tes_category), remove = TRUE, sep = " - ") #combine higher categories into parents
  
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - start simple, set root
  GHGsb1_neg <- data.frame(
    parents =c(rep("", 6)),
    labels =c(paste0(subsectors_neg$tes_subsector)),
    values =as.numeric(c(paste0(subsectors_neg$Temissions)))
  )
  
  ##Add id column
  GHGsb1_neg <- GHGsb1_neg %>%
    mutate(ids = labels) 
  
  ##Now paste together distinct instances into dataframe with hierarchical structure - add on
  GHGsb2_neg <- data.frame(
    parents =c(paste0(cat_neg$tes_subsector), paste0(src_neg$parents)),
    labels =c(paste0(cat_neg$tes_category), paste0(src_neg$source_2)),
    values =as.numeric(c(paste0(cat_neg$Temissions), paste0(src_neg$Temissions)))
  )
  
  ##Add id column                    
  GHGsb2_neg <- GHGsb2_neg %>%
    unite('ids', c(parents, labels), remove = FALSE, sep = " - ") #combine columns for id
  
  ##Combine into one dataframe
  GHGsb_neg <- rbind(GHGsb1_neg, GHGsb2_neg) 
  
  ##Convert negative values to absolute (positive) values
  GHGsb_neg$values <- abs(GHGsb_neg$values)
  
  ## Sunburst Removals with Plotly ####
  sb_neg <- plot_ly(GHGsb_neg,
                    type = 'sunburst',
                    ids = ~ids,
                    labels = ~labels,
                    parents = ~parents,
                    values = ~values,
                    branchvalues = 'total', #parents are totals of children
                    insidetextorientation='radial',
                    maxdepth = 2, #how many levels to show
                    level = "",
                    sort = TRUE #sort by size
                    #width = 800, height = 800 #plot dimensions
  ) %>%
    layout(
      #margin = list(l = 10, r = 10, b = 10, t = 10),
      sunburstcolorway = c(
        "#3CBD9C", #forestry
                 "#F0C954", #grassland
                 "#5D7ED1", #peatland
                 "#C24841", #other
                 "#96BC61",
                 "#F0A7CA", 
                 "#D38744", 
                 "#A780BB"
      ),
      extendsunburstcolors = TRUE
    ) #%>%
    style(sb_neg, hovertemplate="%{label} 
      %{value:,.3r} Mt CO2 equivalents
      %{percentParent:.1%} of %{parent} removals
      <extra></extra>")
  
  #sb_neg
  })

plotly::renderPlotly({produce_rplot()})

```

### Time Series Chart

```{r}

produce_tplot <- shiny::reactive({
  
  #Plot time series ggplot
  Sect_time <- GHG_UK22 %>%
    filter(ghg_grouped %in% input$select_GHG) %>%
    group_by(year, tes_sector) %>%
    summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
    rename("Sector" = "tes_sector") %>% #rename variable
    ggplot(aes(x=as.numeric(year), y=Temissions, color=Sector)) +
    geom_line() +
    xlab("") + ylab("Greenhouse Gas Emissions (Megatonnes CO2 eq.)") +
    ggtitle("Greenhouse Gas Emissions of the UK per Sector Over Time") +
    theme_few() +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = pretty(c(1990:2020), n=6)) +
    scale_color_brewer(palette = "Set2") +
    theme(legend.position="bottom")
    
    #Make interactive with Plotly
    t_int <- plotly::ggplotly(Sect_time) %>% #convert to interactive graph
      layout(legend = list(
        orientation = "h")) #%>%
    style(t_int, hovertemplate="Year: %{x:,.r} #define what hover label says, round x values
          CO2eq: %{y:,.4r} kt") #round y values to 4 sig. figs
    
    #t_int
  
  })


renderPlotly({produce_tplot()})


```

Row {.tabset .tabset-fade}
-------------------------------------
   
### Agriculture

#### Agriculture Emissions by Subsector

```{r}

produce_agplot <- shiny::reactive({

#Bar plot by subsector, stacked by GHG with Plotly
agbar_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  group_by(tes_subsector, ghg, ghg_grouped) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  pivot_wider(names_from = ghg_grouped, values_from=Temissions) %>%
  droplevels() %>%
  plot_ly(x = ~tes_subsector, type = 'bar', y = ~CO2, name = 'CO2', marker = list(color ="#3CBD9C")) %>%
  add_trace(y = ~CH4, name = 'CH4', marker = list(color ="#C24841")) %>%
  add_trace(y = ~N2O, name = 'N2O', marker = list(color ="#D38744")) %>%
  layout(yaxis = list(title = 'Emissions (Mt CO2 eq.)'), 
         barmode = 'stack'
         )

#Plotly customization
style(agbar_int, hovertemplate="Subsector: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agplot()})


```   

#### Agriculture Emissions by Category

```{r}

produce_agcatplot <- shiny::reactive({
  
#Bar plot by category, filled by subsector with Plotly
agcat_int <- GHG_UK22 %>% 
  filter(tes_sector == "Agriculture") %>%
  group_by(tes_subsector, tes_category) %>%
  summarize(Temissions = sum(emissions_mt_co2e, na.rm = TRUE)) %>%
  droplevels() %>%
  plot_ly(x = ~tes_category, type = 'bar', y = ~Temissions, color = ~tes_subsector, colors = colors,
          marker = list(
            width = 1.5)) %>%
  layout(title = 'Agriculture Emissions by TES Subsector and Category',
         yaxis = list(title = 'Emissions (Mt CO2 eq.)'),
         xaxis = list(title = '', tickangle = -90),
         legend = list(x = 0.8, y = 1, orientation = 'v'),
         autosize = F, width = 1000, height = 600,
         autoexpand = T,
         barmode = 'group'
         )

#Plotly customization
style(agcat_int, 
hovertemplate="Category: %{x}
Emissions: %{y:,.4r} Mt CO2 eq.") #define what hover label says, round x values, round y values to 4 sig. figs

})

renderPlotly({produce_agcatplot()})
  
```

### Land Use, Land Use Change, and Forestry

```{r}


```   

### Waste

```{r}


```   

